---
title: "consutling"
author: "Zhihui Zhang 
         Shicong Wang
         Qiannan Shen"
date: "3/28/2022"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r , include = FALSE}
#package we use
library(readr)
library(stringr)
library(ggplot2)
library(dplyr)
library(stringr)
```


```{r , include = FALSE}
#import data
dat <- read_csv('data.csv')
```

## Introduction
The main purpose of this project is evaluating the genetic counseling and germline genetic testing process within the pancreatic adenocarcinoma patient population at Roswell Park Comprehensive Cancer Center. We are mainly interested in investigating the reasons that patients opt out of genetic testing and genetic counseling and the trend of referral rate after the guideline came out. The report can be divided into three parts. The first part will analyze the reason in each step on why people drop out. The second part will focus on the overall trend of the referral rate for 30 months. The third part will compare the referral rate before and after the genetic center offered telehealth and help to answer the question whether people prefer telehealth to in-person meeting. 


## Data Processing 

```{r , include = FALSE}
#second opinion tag 
dat$second_opinion <- str_detect(dat$`Reason No Referral`, pattern = 'second opinion') | str_detect(dat$`Somatic Testing
(y,n,other)`, pattern = 'second opinion')

#number of patients with second opinion
sum(dat$second_opinion)

#the data we will use for the further analysis(exclude patients with second opinion)
dat_1 <- dat[dat$second_opinion == FALSE,]

#referral rate = ordered test / total number
#re-encoding 
dat_1$test <- ifelse(str_detect(dat_1$`Testing Ordered`, pattern = 'yes'), 1, 0)

#referral rate through time 
dat_1$`Initial Appt (#1-30)`[dat_1$`Initial Appt (#1-30)`> 15 & dat_1$`Initial Appt (#1-30)`< 16]  <- 15
dat_1$`Initial Appt (#1-30)`[dat_1$`Initial Appt (#1-30)`> 22 & dat_1$`Initial Appt (#1-30)`< 23]  <- 22

```


## I.Referral rate and its analysis

The overall referral rate is 0.3548387. There are 55 patients ordered the germline genetic test among 155 patients. We will divide the referral process into four steps and analyze the reasons of dropping of during each step.  

### Part.I Genetic referrals placed

```{r,include = FALSE}
data<- dat_1

# probability that place genetic referrals
data$`Genetics Referrals`<- ifelse(data$`# Genetics Referrals` >0, "yes", "no")
#sum(data$`Genetics Referrals` == "yes") / length(data$`Genetics Referrals`)
#sum(data$`Genetics Referrals` == "no") / length(data$`Genetics Referrals`)

```

|Options          |Number of patients|  Ratio |
|----------------:|-----------------:|-------:|
|Yes - Acceptance |               111|   0.716| 
|No               |                44|   0.284|  
|Total            |               155|   -    | 

### Part II. Genetic meeting scheduled
```{r,include = FALSE}
# probability that schedule genetic meeting
data$`Genetics Appt Scheduled`<- ifelse(data$`Genetics Appt Scheduled` == "yes", "yes" , "no")
data$`Genetics Appt Scheduled2`<- ifelse(data$`Genetics Appt Scheduled` == "yes", "yes" ,
                                      ifelse(data$`Genetics Appt Scheduled` =="no", "no", "other"))
#sum(data$`Genetics Appt Scheduled` == "yes") / sum(data$`Genetics Referrals` == "yes")
```

|Options          |Number of patients|  Ratio|
|----------------:|-----------------:|-------:|
|Yes - Acceptance |                81|   0.730| 
|No               |                30|   0.270|  
|Total            |               111|   -    | 

### Part III. Genetic meeting attended
```{r,include = FALSE}
data$`Attended Appt`[data$`Attended Appt` == "yes x2"] <- "yes"
data$`Attended Appt`[data$`Attended Appt` == "???" | data$`Attended Appt` == "n/a"] <- "NA"

data$`Attended Appt`<- ifelse(data$`Attended Appt` == "yes", "yes", "no")
data$`Attended Appt2`<- ifelse(data$`Attended Appt` == "yes", "yes", 
                            ifelse(data$`Attended Appt` == "no", "no", "other"))
# probability that attend the appointment
sum(data$`Attended Appt` %in% 'yes') / sum(data$`Genetics Appt Scheduled` %in% 'yes')

```

|Options          |Number of patients|   Ratio|
|----------------:|-----------------:|-------:|
|Yes - Acceptance |                59|   0.728| 
|No               |                22|   0.272|  
|Total            |                81|   -    | 

### Part IV. Genetic counseling ordered 
```{r,include = FALSE}
data$`Testing Ordered`[data$`Testing Ordered` =="yes on second"]<- "yes"
data$`Testing Ordered`[data$`Testing Ordered` == "NO - PREVIOUSLY COMPLETED OUTSIDE" | data$`Testing Ordered` == "other - discussed DNA banking b/c hospice"]<- "no"
data$`Testing Ordered`[data$`Testing Ordered` =="??" | data$`Testing Ordered` =="n/a" ]<- "na"
data$`Testing Ordered`<- ifelse(data$`Testing Ordered` == "yes", "yes", "no")
data$`Testing Ordered2`<- ifelse(data$`Testing Ordered` == "yes", "yes", 
                               ifelse(data$`Testing Ordered` == "no","no","other"))
# probability that order genetic counseling 
 sum(data$`Testing Ordered` %in% 'yes') / sum(data$`Attended Appt` %in% 'yes')
```

|Options          |Number of patients|   Ratio|
|----------------:|-----------------:|-------:|
|Yes - Acceptance |                55|   0.932| 
|No               |                 4|   0.068|  
|Total            |                59|   -    | 


### V. Proportions of individuals who proceed through each step of the process:
The following Sankey plot illustrates how many individuals succeed (currently labeled “Accept”) at each step of the genetic counseling process. 

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(webshot)

#install phantom:
webshot::install_phantomjs()
# Make a webshot in pdf : high quality but can not choose printed zone
webshot("file:///Users/shicongwang/Desktop/genetic_counseling/Untitled.html" , "output.pdf", delay = 0.2)

```


\newpage

## II. Other Analysis

### I.Cancer Stage and Genetic Counseling Referral Rate

We thought that the stage of cancer might explain some of the variation in referral rate. The following stacked bar plot serves as visualization to the proportions of individuals who ordered testing given that they were diagnosed with certain stages of cancer. We only selected the patients who finally ordered genetic counseling, and labeled the numbers and percentages on each bar.The percentage can be calculated by:
$$percentage = \frac{number \ of \ patients \ ordered \ testing}{number \ of \ patients\ in \ each \ stage}$$
For instance, the total number of patients in fourth stage is 121, and 41 of them ordered genetic counseling, so the percentage is supposed to be 41/121=33.88%.

```{r echo=FALSE, warning=FALSE, message=FALSE}
data$`Testing Ordered` <- ifelse(str_detect(data$`Testing Ordered`, 'yes'), 'yes', 'no')
df5<- data %>%
  count(`Testing Ordered`,`Stage at Dx (#0-4)`) %>%
  arrange(`Testing Ordered`, desc(`Stage at Dx (#0-4)`)) %>% # Rearranging in stacking order  
  group_by(`Stage at Dx (#0-4)`) %>%
  mutate(Freq2 = cumsum(n), # Calculating position of stacked Freq
         prop = n/sum(n),
         percentage = 100* n/sum(n))  # Calculating proportion of Freq

df5$label = paste0(sprintf("%.2f", df5$percentage), "%")
df5$sd<- sqrt( df5$prop*(1-df5$prop) / df5$Freq2)

df_5<- df5 %>% filter(`Testing Ordered`== "yes")

ggplot(df_5) +
    geom_bar( aes(x=`Stage at Dx (#0-4)`, y=prop), stat="identity", fill="skyblue", alpha=0.7) +
    geom_errorbar(aes(x=`Stage at Dx (#0-4)`, ymin=prop-sd, ymax=prop+sd), width=0.4, colour="orange", alpha=0.9, size=1.3)+
    geom_text(aes(x=`Stage at Dx (#0-4)`, y=prop,label = n), position = position_stack(vjust = 0.5), size = 4, color="white")+
    geom_text(aes(x=`Stage at Dx (#0-4)`, y=prop,label = label), position = position_stack(vjust = 0.3), size = 4, color="white")+
    scale_y_continuous(labels = scales::percent)+
    ylab("Percentage")+
    xlab("Stages")
      

```

### II. ECOG and Referral Rate

ECOG is an indicator of the wellness level of the patients. In the following section, we want to investigate the relationship between ECOG of patients and the referral rate. As a result, we use the same method as the previous analysis.The stacked bar plot presents the proportions of individuals who ordered testing based on the patient’s level of functioning in terms of their ability to care for themselves (ECOG Status).

```{r echo=FALSE, warning=FALSE, message=FALSE}

df6<- data %>%
  count(`Testing Ordered`,`ECOG at Initial (#0-4)`) %>%
  arrange(`Testing Ordered`, desc(`ECOG at Initial (#0-4)`)) %>% # Rearranging in stacking order  
  group_by(`ECOG at Initial (#0-4)`) %>%
  mutate(Freq2 = cumsum(n), # Calculating position of stacked Freq
         prop = n/sum(n),
         percentage = 100* n/sum(n))  # Calculating proportion of Freq

df6$label = paste0(sprintf("%.2f", df6$percentage), "%")

df6$sd<- sqrt( df6$prop*(1-df6$prop) / df6$Freq2)

df_6<- df6 %>% filter(`Testing Ordered`== "yes")


#ggplot(df6, aes(x = `ECOG at Initial (#0-4)`, y = n, fill = `Testing Ordered`)) +
   #geom_bar(position = position_stack(), stat = "identity") +
   #geom_text(aes(label = label), position = position_stack(vjust = 0.5), size = 3, color="white")+
   #ylab("Frequency")+
   #xlab("ECOG")

#ggplot(df6, aes(x = `ECOG at Initial (#0-4)`, y = n, fill = `Testing Ordered`)) +
   #geom_bar(position = "dodge", stat = "identity") +
   #geom_text(aes(label=n), position=position_dodge(width=0.9), vjust=-0.25)+
   #ylab("Frequency")+
   #xlab("ECOG")

ggplot(df_6) +
    geom_bar( aes(x=`ECOG at Initial (#0-4)`, y=prop), stat="identity", fill="skyblue", alpha=0.7) +
    geom_errorbar(aes(x=`ECOG at Initial (#0-4)`, ymin=prop-sd, ymax=prop+sd), width=0.4, colour="orange", alpha=0.9, size=1.3)+
    geom_text(aes(x=`ECOG at Initial (#0-4)`, y=prop,label = n), position = position_stack(vjust = 0.5), size = 4, color="white")+
    geom_text(aes(x=`ECOG at Initial (#0-4)`, y=prop,label = label), position = position_stack(vjust = 0.3), size = 4, color="white")+
    scale_y_continuous(labels = scales::percent)+
    ylab("Percentage")+
    xlab("ECOG Status")
```


## III. Change of referral rate

In this section, we visualize the trends of referral rate using both 3 months and 6 months as a period. We can learn whether the referral rate is increasing or not through the time. 

```{r, include = FALSE}
#referral rate through time 
dat_2 <- dat_1
dat_2$`Initial Appt (#1-30)`[dat_2$`Initial Appt (#1-30)`> 15 & dat_2$`Initial Appt (#1-30)`< 16]  <- 15
dat_2$`Initial Appt (#1-30)`[dat_2$`Initial Appt (#1-30)`> 22 & dat_2$`Initial Appt (#1-30)`< 23]  <- 22

dat_2$cut_3 <- cut(dat_2$`Initial Appt (#1-30)`, c(1, seq(3, 30, by=3)), include.lowest = T) 
dat_2$cut_6 <- cut(dat_2$`Initial Appt (#1-30)`, c(1, seq(6, 30, by=6)), include.lowest = T) 

```


```{r echo=FALSE, warning=FALSE, message=FALSE}
dat1 <- dat_2 %>% group_by(`Initial Appt (#1-30)`,cut_3) %>%
  summarize(referral_rate= sum(test)/ n(), cut_3) %>% distinct()

dat2 <- dat1 %>% group_by(cut_3) %>%
 summarise(sd = sd(referral_rate),
          referral_rate = mean(referral_rate))


ggplot(dat1 , aes(x = cut_3, y = referral_rate)) +
  geom_jitter(position = position_jitter(0.15), color = "darkgray") + 
  geom_line(aes(group = 1), color="#69b3a2", data = dat2) +
  geom_errorbar(data = dat2, aes(ymin = referral_rate-sd, ymax =referral_rate+sd), width = 0.2)  +
  geom_point(data = dat2, size = 2) + 
  labs(title="Referral rate change per 3 month", x="Month", y="Referral rate")
```

The plot above shows us the trend of referral rate using 3 month as a period.We jitter the points to avoid overlaps. The plot above does not indicate there is any increasing or decreasing trend. 

```{r echo=FALSE, message=FALSE}
dat_2$cut_6 <- cut(dat_2$`Initial Appt (#1-30)`, c(1, seq(6, 30, by=6)), include.lowest = T) 

dat1 <- dat_2 %>% group_by(`Initial Appt (#1-30)`,cut_6) %>%
  summarize(referral_rate= sum(test)/ n(), cut_6) %>% distinct()

dat2 <- dat1 %>% group_by(cut_6) %>%
 summarise(sd = sd(referral_rate),
          referral_rate = mean(referral_rate))


ggplot(dat1 , aes(x = cut_6, y = referral_rate)) +
  geom_jitter(position = position_jitter(0.2), color = "darkgray") + 
  geom_line(aes(group = 1), data = dat2,  color="#69b3a2") +
  geom_errorbar(data = dat2, aes(ymin = referral_rate-sd, ymax =referral_rate+sd), width = 0.2)  +
  geom_point(data = dat2, size = 2)+ 
   labs(title="Referral rate change per 6 month", x="Month", y="Referral rate")
```

The visualization above tells us the change of the referral rate per 6 month. Since there are a lot of variations of the referral rate in each time interval, we can not conclude that there is an obvious increasing relationship between time and referral rate. 



\newpage

## III.Virtual/Teleheath Impact

In this part, we want to investigate if there is any difference of the referral rate before and after telehealth is provided as an preference options for the patients. 

|Period|Total number of patients|Number of patients who placed a referral|Number of Patients who scheduled a meeting|Number of patients who attended the meeting|Number of Patients who ordered the test|Referral rate|
|----------:|----------:|----------:|----------:|----------:|----------:|----------:|
|Before     |        120|         82|         55|         40|         37|     0.3083|
|After      |         35|         29|         26|         18|         17|     0.4857|
| Telehealth|     -     |     -     |         17|         14|         13|     -     |
|  In-person|     -     |     -     |          5|          4|          4|     -     |
|         NA|     -     |     -     |          4|          0|          0|     -     |
|Total      |        155|        111|         81|         58|         54|     0.3484| 

Qeustion: 


\newpage

## IV.Analysis on second opinion patients


Among 65 patients who had second opinion in the data, 3 of them still ordered the germline genetic testing.

```{r, include = FALSE}
dat_3 <- dat[dat$second_opinion == TRUE, ]
nrow(dat_3)
nrow(dat_3 %>% filter(`# Genetics Referrals`>=1))
nrow(dat_3 %>% filter(`# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'))
nrow(dat_3 %>% filter(`# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'))
sum(dat_3$`Testing Ordered` == 'yes')
```
  

|Period|Total number of patients|Number of patients who placed a referral|Number of Patients who scheduled a meeting|Number of patients who attended the meeting|Number of Patients who ordered the test|Referral rate|
|----------:|----------:|----------:|----------:|----------:|----------:|----------:|
|Before     |         65|         11|          8|          6|          3|    0.04615|


\newpage

## V.Reasons for failure

### I. Reasons that cause the patients with no second opinion fail to place the referral

We next explore the reason that causes the patients to fail to schedule genetic meetings and the bar plot shown below. The graph shows that the main reason is ‘pt deferred/declined’ which is about 33%. The second most important reason should be ‘no response by call/mail’ that about 9 patients (27%) fail due to this reason. The resting patients fail to schedule genetic meetings caused by ‘never contacted’ (9%), ‘never schd’ (9%), and ‘not addressed’ (15%). 


```{r echo=FALSE, warning=FALSE, message=FALSE}
dat <- read.csv("data.csv")
df1 <- dat[c(12,14)]

df1$Re1.second <- ifelse(str_detect(df1[,2], pattern = 'second'), 1, 0)
df10 <- df1 %>% filter(Reason.No.Referral != "n/a")
df11 <- df1 %>% filter(Reason.No.Referral != "n/a") %>% filter(Re1.second == 0)

df11$Re1.confirm <- ifelse(str_detect(df11[,2], pattern = 'confirm dx'), 1, 0)
df11$Re1.trial <- ifelse(str_detect(df11[,2], pattern = 'trial'), 1, 0)
df11$Re1.palliative <- ifelse(str_detect(df11[,2], pattern = 'palliative'), 1, 0)
df11$Re1.hospice <- ifelse(str_detect(df11[,2], pattern = 'hospice'), 1, 0)
df11$Re1.testing <- ifelse(str_detect(df11[,2], pattern = 'testing'), 1, 0)

count1 <- c(sum(df11[4]=="1"),sum(df11[5]=="1"),
            sum(df11[6]=="1"),sum(df11[7]=="1"),sum(df11[8]=="1"))
reasons1 <- c("confirm dx","trial","palliative","hospice","testing")
result1 <- data.frame(reasons1, count1)

ggplot(data=result1, aes(x=reasons1, y=count1)) +
  geom_bar(stat="identity", col = "skyblue", fill = "skyblue")+
  geom_text(aes(label=count1), vjust=1.6, color="white", size=3.5)+
  theme_bw()+
  ggtitle("Reasons that cause the patients with no second opinion fail to place the referral")
```
Encoding: 
confirm: confirm dx then chose care at ECMC, workup/confirm dx;

trial: phase III trial, trial screen;

palliative: palliative/ supportive, recommended supportive/ palliative care, other - palliative/ supportive;

testing: other - prior testing completed at Roswell (2017), other - prior testing previously completed outside;

hospice: hospice, then inpatient then enrolled in hospice, other - hospice

\newpage

### II. Reasons that cause the patients fail to schedule genetic meeting

We next explore the reason that causes the patients to fail to schedule genetic meetings and the bar plot shown below. The graph shows that the main reason is ‘no response by call/mail’ that about 9 patients (27%) fail due to this reason. The resting patients fail to schedule genetic meetings caused by ‘never contacted’ (9%), ‘never schd’ (9%), ‘not addressed’ (15%), 'passed' (15%'), and 'pt declined' (18%). 


```{r echo=FALSE, warning=FALSE, message=FALSE}
df2 <- dat[c(18,19)]

df2$pt <- ifelse(str_detect(df2[,2], pattern = 'pt'), 1, 0)
df2$no <- ifelse(str_detect(df2[,2], pattern = 'no'), 1, 0)
df2$neverc <- ifelse(str_detect(df2[,2], pattern = 'never contacted'), 1, 0)
df2$nevers <- ifelse(str_detect(df2[,2], pattern = 'never schd'), 1, 0)
df2$address <- ifelse(str_detect(df2[,2], pattern = 'address'), 1, 0)
df2$pass <- ifelse(str_detect(df2[,2], pattern = 'pass'), 1, 0)

count2 <- c(sum(df2[3]=="1")-sum(df2[8]=="1"),sum(df2[4]=="1")-7,sum(df2[5]=="1"),
            sum(df2[6]=="1"),sum(df2[7]=="1"),sum(df2[8]=="1"))
reasons2 <- c("pt deferred/ declined ","no responses","never contancted", 
              "never schd", "not addressed", "passed")
result2<- data.frame(reasons2,count2)

ggplot(data=result2, aes(x=reasons2, y=count2)) +
  geom_bar(stat="identity", col = "pink", fill = "pink")+
  geom_text(aes(label=count2), vjust=1.6, color="white", size=3.5)+
  theme_bw() +
  ggtitle("Reasons that cause the patients fail to schedule genetic meeting")
```
Encoding: 
pt declined: pt deferred/not interested, pt deferred/ declined, other - pt declined x2;

No response/mail: other - no response to calls/VM or mailed letter, other - no call back, inpatient to inpatient hospice;

Never schd: no never schd, never seen;

Not addressed: referral not addressed - active;

Never contacted: never contacted - d/c via pt discharge;

passed: pt passed before referral reviewed, pt passed before appt offered, never contacted - d/c via pt discharge;

\newpage

## Appendix
```{r include = FALSE}
#overall referral rate
sum(dat_1$test)/length(dat_1$test)

#referral rate before and after two options
#before
sum(dat_1$test[dat_1$`Initial Appt (#1-30)`<=22.2])/length(dat_1$test[dat_1$`Initial Appt (#1-30)`<=22.2])
#after
sum(dat$test[dat_1$`Initial Appt (#1-30)`>22.2])/length(dat_1$test[dat_1$`Initial Appt (#1-30)`>22.2])

```

```{r include = FALSE}
#before and after two options

nrow(dat_1 %>% filter(`Initial Appt (#1-30)`<= 22.2)) #total patients before
nrow(dat_1 %>% filter(`Initial Appt (#1-30)` > 22.2)) #total patients after

nrow(dat_1 %>% filter(`Initial Appt (#1-30)`<= 22.2& `# Genetics Referrals`>=1)) #total patients who had referrals before 
nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2& `# Genetics Referrals`>=1) ) #total patients who had referrals after


nrow(dat_1 %>% filter(`Initial Appt (#1-30)`<= 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'))#total patients who scheduled the appt before 
nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'))#total patients who scheduled the appt after

nrow(dat_1 %>% filter(`Initial Appt (#1-30)`<= 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'))#total patients who scheduled the appt before
nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'))#total patients who scheduled the appt after

nrow(dat_1 %>% filter(`Initial Appt (#1-30)`<= 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'& `Testing Ordered` == 'yes'))#total patients who ordered the test before 
nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'& `Testing Ordered` == 'yes'))#total patients who ordered the test after 


nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Virtual Offered` %in% c('yes - preference', 'yes - preference (distance)', 'yes'))) # patients who prefered virtual and scheduled the test
nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Virtual Offered` %in% c('no', 'no - preference'))) # patients who prefered person and scheduled the test

table(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes') %>% select(`Virtual Offered`))

nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes' & `Virtual Offered` %in% c('yes - preference', 'yes - preference (distance)', 'yes'))) # patients who prefered virtual and attended the test
nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'& `Virtual Offered` %in% c('no', 'no - preference'))) # patients who prefered person and attended the test

table(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes') %>% select(`Virtual Offered`))


table(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'&`Testing Ordered` == 'yes') %>% select(`Virtual Offered`))

```
