---
title: "Analysis of Genetic Counseling among Pancreatic Cancer Patients"
author: "Zhihui Zhang, Shicong Wang, Qiannan Shen"
date: "04/01/2022"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r , include = FALSE}
#package we use
library(readr)
library(stringr)
library(ggplot2)
library(dplyr)
library(stringr)
library(ggpubr)
library(webshot)
```


```{r , include = FALSE}
#import data
dat <- read_csv('data.csv')
```

## Introduction

The main purpose of this project is to evaluate the genetic counseling and germline genetic testing process within the pancreatic adenocarcinoma patient population at Roswell Park Comprehensive Cancer Center. We are mainly interested in investigating the reasons that patients opt out of genetic testing and genetic counseling and the trend of referral rate after the guideline came out. The report can be divided into five parts. The first part will analyze the reasons in each step why people drop out. We will uncover the relationship between referral rate and other factors including disease stage and ECOG score. The third part will focus on the overall trend of the referral rate for 30 months. The fourth part will compare the referral rate before and after the genetic center offered telehealth and help to answer the question of whether people prefer telehealth to in-person meetings. The last part is the additional analysis of people with second opinions as well as an exploration of the dominant reasons for dropping-out individuals.

## Data Processing 

The dataset includes 220 pancreatic patients' medical records for 30 months. The dataset contains personal information of patients, medical records as well as genetic counseling records. There are many missing values in the data. We fill in some missing values based on our understanding of context. For example, the entry will be missing for the later part of genetic counseling if the patient does not have a referral placed. In further analysis, we also divide some specific reasons into groups. The detailed encoding information will be shown in the Appendix. Also, except for the analysis of second opinion patients, we only include 155 patients who do not have second opinions in either somatic testing or the reason no referral column in the report. 

```{r , include = FALSE}
#second opinion tag 
dat$second_opinion <- str_detect(dat$`Reason No Referral`, pattern = 'second opinion') | str_detect(dat$`Somatic Testing
(y,n,other)`, pattern = 'second opinion')

#number of patients with second opinion
sum(dat$second_opinion)

#the data we will use for the further analysis(exclude patients with second opinion)
dat_1 <- dat[dat$second_opinion == FALSE,]

#referral rate = ordered test / total number
#re-encoding 
dat_1$test <- ifelse(str_detect(dat_1$`Testing Ordered`, pattern = 'yes'), 1, 0)

#referral rate through time 
dat_1$`Initial Appt (#1-30)`[dat_1$`Initial Appt (#1-30)`> 15 & dat_1$`Initial Appt (#1-30)`< 16]  <- 15
dat_1$`Initial Appt (#1-30)`[dat_1$`Initial Appt (#1-30)`> 22 & dat_1$`Initial Appt (#1-30)`< 23]  <- 22

```

## Summary of Analysis

## I. Analysis on Referral Rate 

The overall referral rate is 35.5%. There were 55 patients who ordered the germline genetic test among 155 patients. We will divide the referral process into four steps and analyze the reasons for dropping off during each step in later analysis. 

$$ Referral \ rate = \frac{number \ of \ patients \ ordered \ testing}{total \ number \ of \ patients}$$

\newpage

### Part.I Genetic referrals placed

The first acceptance ratio refers to the number of patients whose genetic referrals were placed by doctors over the total amount of patients. 

```{r,include = FALSE}
data<- dat_1

# probability that place genetic referrals
data$`Genetics Referrals`<- ifelse(data$`# Genetics Referrals` >0, "yes", "no")
#sum(data$`Genetics Referrals` == "yes") / length(data$`Genetics Referrals`)
#sum(data$`Genetics Referrals` == "no") / length(data$`Genetics Referrals`)

```

|Options          |Number of patients|  Ratio |
|----------------:|-----------------:|-------:|
|Yes - Acceptance |               111|   0.716| 
|No               |                44|   0.284|  
|Total            |               155|   -    | 


### Part II. Genetic meetings scheduled

The second acceptance ratio is calculated based on the previous step, which implies the proportion of patients who scheduled the testing by the patients who accepted the first step.

```{r,include = FALSE}
# probability that schedule genetic meeting
data$`Genetics Appt Scheduled`<- ifelse(data$`Genetics Appt Scheduled` == "yes", "yes" , "no")
data$`Genetics Appt Scheduled2`<- ifelse(data$`Genetics Appt Scheduled` == "yes", "yes" ,
                                      ifelse(data$`Genetics Appt Scheduled` =="no", "no", "other"))
#sum(data$`Genetics Appt Scheduled` == "yes") / sum(data$`Genetics Referrals` == "yes")
```

|Options          |Number of patients|  Ratio|
|----------------:|-----------------:|-------:|
|Yes - Acceptance |                81|   0.730| 
|No               |                30|   0.270|  
|Total            |               111|   -    | 


### Part III. Genetic meetings attended

The calculation of the third ratio is the same as the second one, which represents the number of patients attending a genetic meeting over the number of the acceptance group in the second step.

```{r,include = FALSE}
data$`Attended Appt`[data$`Attended Appt` == "yes x2"] <- "yes"
data$`Attended Appt`[data$`Attended Appt` == "???" | data$`Attended Appt` == "n/a"] <- "NA"

data$`Attended Appt`<- ifelse(data$`Attended Appt` == "yes", "yes", "no")
data$`Attended Appt2`<- ifelse(data$`Attended Appt` == "yes", "yes", 
                            ifelse(data$`Attended Appt` == "no", "no", "other"))
# probability that attend the appointment
sum(data$`Attended Appt` %in% 'yes') / sum(data$`Genetics Appt Scheduled` %in% 'yes')

```

|Options          |Number of patients|   Ratio|
|----------------:|-----------------:|-------:|
|Yes - Acceptance |                59|   0.728| 
|No               |                22|   0.272|  
|Total            |                81|   -    | 

### Part IV. Genetic counselings ordered 

In the table below, we calculate the proportion of patients who ordered testing given the third acceptance group.

```{r,include = FALSE}
data$`Testing Ordered`[data$`Testing Ordered` =="yes on second"]<- "yes"
data$`Testing Ordered`[data$`Testing Ordered` == "NO - PREVIOUSLY COMPLETED OUTSIDE" | data$`Testing Ordered` == "other - discussed DNA banking b/c hospice"]<- "no"
data$`Testing Ordered`[data$`Testing Ordered` =="??" | data$`Testing Ordered` =="n/a" ]<- "na"
data$`Testing Ordered`<- ifelse(data$`Testing Ordered` == "yes", "yes", "no")
data$`Testing Ordered2`<- ifelse(data$`Testing Ordered` == "yes", "yes", 
                               ifelse(data$`Testing Ordered` == "no","no","other"))
# probability that order genetic counseling 
 sum(data$`Testing Ordered` %in% 'yes') / sum(data$`Attended Appt` %in% 'yes')
```

|Options          |Number of patients|   Ratio|
|----------------:|-----------------:|-------:|
|Yes - Acceptance |                55|   0.932| 
|No               |                 4|   0.068|  
|Total            |                59|   -    | 


\newpage

### Part V. Proportions of individuals who proceed through each step of the process

The following Sankey plot illustrates how many individuals succeed (currently labeled “Accept”) at each step of the genetic counseling process. 

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.cap= 'Proportions of Patients in each Step'}


#install phantom:
webshot::install_phantomjs()
# Make a webshot in pdf : high quality but can not choose printed zone
webshot("sankey_plot.html" , file = "output.pdf", delay = 0.2)
```


\newpage

## II. Analysis of Referral Rate with Other Factors

### Part I.Cancer Stage and Referral Rate

We think that the stage of cancer might explain some of the variations in referral rate. The following stacked bar plot serves as a visualization of the proportions of individuals who ordered testing given that they were diagnosed with certain stages of cancer. We only selected the patients who finally ordered genetic counseling and labeled the numbers and percentages on each bar.The percentage can be calculated by:
$$percentage = \frac{number \ of \ patients \ ordered \ testing}{number \ of \ patients\ in \ each \ stage}$$
For instance, the total number of patients in fourth stage is 121, and 41 of them have ordered genetic counseling, so the percentage is supposed to be $\frac{41}{121}=33.88\%$.

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.cap= 'Proportion of Patients Who Ordered Tests in each Cancer Stage'}
data$`Testing Ordered` <- ifelse(str_detect(data$`Testing Ordered`, 'yes'), 'yes', 'no')
df5<- data %>%
  count(`Testing Ordered`,`Stage at Dx (#0-4)`) %>%
  arrange(`Testing Ordered`, desc(`Stage at Dx (#0-4)`)) %>% # Rearranging in stacking order  
  group_by(`Stage at Dx (#0-4)`) %>%
  mutate(Freq2 = cumsum(n), # Calculating position of stacked Freq
         prop = n/sum(n),
         percentage = 100* n/sum(n))  # Calculating proportion of Freq

df5$label = paste0(sprintf("%.2f", df5$percentage), "%")
df5$sd<- sqrt( df5$prop*(1-df5$prop) / df5$Freq2)

df_5<- df5 %>% filter(`Testing Ordered`== "yes")

ggplot(df_5) +
    geom_bar( aes(x=`Stage at Dx (#0-4)`, y=prop), stat="identity", fill="skyblue", alpha=0.7) +
    geom_errorbar(aes(x=`Stage at Dx (#0-4)`, ymin=prop-sd, ymax=prop+sd), width=0.4, colour="orange", alpha=0.9, size=1.3)+
    geom_text(aes(x=`Stage at Dx (#0-4)`, y=prop,label = n), position = position_stack(vjust = 0.5), size = 4, color="white")+
    geom_text(aes(x=`Stage at Dx (#0-4)`, y=prop,label = label), position = position_stack(vjust = 0.3), size = 4, color="white")+
    scale_y_continuous(labels = scales::percent)+
    ylab("Percentage")+
    xlab("Stages")
      

```

\newpage

### Part II. ECOG and Referral Rate

ECOG is an indicator of the wellness level of the patients. In the following section, we want to investigate the relationship between the ECOG of patients and the referral rate. As a result, we use the same method as the previous analysis. The stacked bar plot presents the proportions of individuals who ordered testing based on the patient’s level of functioning in terms of their ability to care for themselves (ECOG Status).

```{r echo = FALSE, warning = FALSE, message = FALSE, fig.cap= 'Proportions of Patients Who Ordered tests in each ECOG Level'}

df6 <- data %>%
  count(`Testing Ordered`,`ECOG at Initial (#0-4)`) %>%
  arrange(`Testing Ordered`, desc(`ECOG at Initial (#0-4)`)) %>% # Rearranging in stacking order  
  group_by(`ECOG at Initial (#0-4)`) %>%
  mutate(Freq2 = cumsum(n), # Calculating position of stacked Freq
         prop = n/sum(n),
         percentage = 100* n/sum(n))  # Calculating proportion of Freq

df6$label = paste0(sprintf("%.2f", df6$percentage), "%")

df6$sd<- sqrt( df6$prop*(1-df6$prop) / df6$Freq2)

df_6<- df6 %>% filter(`Testing Ordered`== "yes")


#ggplot(df6, aes(x = `ECOG at Initial (#0-4)`, y = n, fill = `Testing Ordered`)) +
   #geom_bar(position = position_stack(), stat = "identity") +
   #geom_text(aes(label = label), position = position_stack(vjust = 0.5), size = 3, color="white")+
   #ylab("Frequency")+
   #xlab("ECOG")

#ggplot(df6, aes(x = `ECOG at Initial (#0-4)`, y = n, fill = `Testing Ordered`)) +
   #geom_bar(position = "dodge", stat = "identity") +
   #geom_text(aes(label=n), position=position_dodge(width=0.9), vjust=-0.25)+
   #ylab("Frequency")+
   #xlab("ECOG")

ggplot(df_6) +
    geom_bar( aes(x=`ECOG at Initial (#0-4)`, y=prop), stat="identity", fill="skyblue", alpha=0.7) +
    geom_errorbar(aes(x=`ECOG at Initial (#0-4)`, ymin=prop-sd, ymax=prop+sd), width=0.4, colour="orange", alpha=0.9, size=1.3)+
    geom_text(aes(x=`ECOG at Initial (#0-4)`, y=prop,label = n), position = position_stack(vjust = 0.5), size = 4, color="white")+
    geom_text(aes(x=`ECOG at Initial (#0-4)`, y=prop,label = label), position = position_stack(vjust = 0.3), size = 4, color="white")+
    scale_y_continuous(labels = scales::percent, limits = c(0,1))+
    ylab("Percentage")+
    xlab("ECOG Status") 
```

\newpage

## III. Change of Referral Rate

In this section, we visualize the trends of referral rates using both three months and six months as a period. We can learn whether the referral rate is increasing or not through time from the figures.

```{r, include = FALSE}
#referral rate through time 
dat_2 <- dat_1
dat_2$`Initial Appt (#1-30)`[dat_2$`Initial Appt (#1-30)`> 15 & dat_2$`Initial Appt (#1-30)`< 16]  <- 15
dat_2$`Initial Appt (#1-30)`[dat_2$`Initial Appt (#1-30)`> 22 & dat_2$`Initial Appt (#1-30)`< 23]  <- 22

dat_2$cut_3 <- cut(dat_2$`Initial Appt (#1-30)`, c(1, seq(3, 30, by=3)), include.lowest = T) 
dat_2$cut_6 <- cut(dat_2$`Initial Appt (#1-30)`, c(1, seq(6, 30, by=6)), include.lowest = T) 

```


```{r echo=FALSE, warning=FALSE, message=FALSE, fig.cap= 'Change of Referral Rate Per 3 Month'}
dat1 <- dat_2 %>% group_by(`Initial Appt (#1-30)`,cut_3) %>%
  summarize(referral_rate= sum(test)/ n(), cut_3) %>% distinct()

dat2 <- dat1 %>% group_by(cut_3) %>%
 summarise(sd = sd(referral_rate),
          referral_rate = mean(referral_rate))


ggplot(dat1 , aes(x = cut_3, y = referral_rate)) +
  geom_jitter(position = position_jitter(0.15), color = "darkgray") + 
  geom_line(aes(group = 1), color="#69b3a2", data = dat2) +
  geom_errorbar(data = dat2, aes(ymin = referral_rate-sd, ymax =referral_rate+sd), width = 0.2)  +
  geom_point(data = dat2, size = 2) + 
  labs(title="Change of Referral Rate Per 3 Month", x="Month", y="Referral rate")
```

The figure above shows us the trend of referral rate using three months as a period. We jitter the points to avoid overlaps. The plot above does not indicate there is any increasing or decreasing trend. 

\newpage

```{r echo=FALSE, message=FALSE, fig.cap= 'Change of Referral Rate Per 6 Month'}
dat_2$cut_6 <- cut(dat_2$`Initial Appt (#1-30)`, c(1, seq(6, 30, by=6)), include.lowest = T) 

dat1 <- dat_2 %>% group_by(`Initial Appt (#1-30)`,cut_6) %>%
  summarize(referral_rate= sum(test)/ n(), cut_6) %>% distinct()

dat2 <- dat1 %>% group_by(cut_6) %>%
 summarise(sd = sd(referral_rate),
          referral_rate = mean(referral_rate))


ggplot(dat1 , aes(x = cut_6, y = referral_rate)) +
  geom_jitter(position = position_jitter(0.2), color = "darkgray") + 
  geom_line(aes(group = 1), data = dat2,  color="#69b3a2") +
  geom_errorbar(data = dat2, aes(ymin = referral_rate-sd, ymax =referral_rate+sd), width = 0.2)  +
  geom_point(data = dat2, size = 2)+ 
   labs(title="Change of Referral Rate Per 6 Month", x="Month", y="Referral rate")
```

The visualization above tells us the change in the referral rate per 6 months. Since there are a lot of variations in the referral rate in each time interval, we can not conclude that there is an obvious increasing relationship between time and referral rate. 



\newpage

## IV.Virtual/Teleheath Impact

In this part, we want to investigate if there is any difference in the referral rate before and after telehealth is provided as a preferred option for the patients. After two options are offered, which corresponds to the 'After' row, the referral rate is 48.6%. The number of patients who preferred telehealth versus those who preferred in-person meetings is the entry in the row 'Telehealth' and 'In-person.

|Period|Total number of patients|Number of patients who placed a referral|Number of Patients who scheduled a meeting|Number of patients who attended the meeting|Number of Patients who ordered the test|Referral rate|
|----------:|----------:|----------:|----------:|----------:|----------:|----------:|
|Before     |        120|         82|         55|         40|         37|     0.308 |
|After      |         35|         29|         26|         18|         17|     0.486 |
| Telehealth|     -     |     -     |         17|         14|         13|     -     |
|  In-person|     -     |     -     |          5|          4|          4|     -     |
|         NA|     -     |     -     |          4|          0|          0|     -     |
|Total      |        155|        111|         81|         58|         54|     0.348 | 



## V. Further Analysis

### Part I.Analysis on 'Second Opinion' patients

Among 65 patients who had 'second opinion' in the data, 3 of them still ordered the germline genetic testings. And the referral rate among these patients is 4.62%.

```{r, include = FALSE}
dat_3 <- dat[dat$second_opinion == TRUE, ]
nrow(dat_3)
nrow(dat_3 %>% filter(`# Genetics Referrals`>=1))
nrow(dat_3 %>% filter(`# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'))
nrow(dat_3 %>% filter(`# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'))
sum(dat_3$`Testing Ordered` == 'yes')
```
  

|Period|Total number of patients|Number of patients who placed a referral|Number of Patients who scheduled a meeting|Number of patients who attended the meeting|Number of Patients who ordered the test|Referral rate|
|----------:|----------:|----------:|----------:|----------:|----------:|----------:|
|Before     |         65|         11|          8|          6|          3|    0.0462|

\newpage

### Part II.Reasons for Failures

#### 1.Reasons that cause the patients with no second opinion fail to place the referrals
\
\
The following bar plots provide simple counts of the number of individuals who listed certain reasons for failing to schedule genetic meetings or place the referral. We firstly separate the patients who fail to place the referrals into two subsets based on whether the failure reason is caused by the second opinion. As result has shown, the number of patients that have a second opinion is 53 and the resting is 48. Then we further explore the reasons for the subset where patients do not contain 'second opinion'. The bar plot below illustrates the reason that causes patients with no second opinion to fail to place the referral. The main reason is 'hospice' (31%). And there are about 8 patients (17%) who fail because of 'screen trials/phase III trials'. The resting patients who failed to place the referrals are caused by  'confirm dx', 'palliative', and 'testing' (around 10%). 

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.cap = 'Reasons for Patients without Referral Placed'}
dat <- read.csv("data.csv")
df1 <- dat[c(12,14)]

df1$Re1.second <- ifelse(str_detect(df1[,2], pattern = 'second'), 1, 0)
df10 <- df1 %>% filter(Reason.No.Referral != "n/a")
df11 <- df1 %>% filter(Reason.No.Referral != "n/a") %>% filter(Re1.second == 0)

df11$Re1.confirm <- ifelse(str_detect(df11[,2], pattern = 'confirm dx'), 1, 0)
df11$Re1.trial <- ifelse(str_detect(df11[,2], pattern = 'trial'), 1, 0)
df11$Re1.palliative <- ifelse(str_detect(df11[,2], pattern = 'palliative'), 1, 0)
df11$Re1.hospice <- ifelse(str_detect(df11[,2], pattern = 'hospice'), 1, 0)
df11$Re1.testing <- ifelse(str_detect(df11[,2], pattern = 'testing'), 1, 0)

count1 <- c(sum(df11[4]=="1"),sum(df11[5]=="1"),
            sum(df11[6]=="1"),sum(df11[7]=="1"),sum(df11[8]=="1"))
reasons1 <- c("confirm dx","trial","palliative","hospice","testing")
result1 <- data.frame(reasons1, count1)

ggplot(data=result1, aes(x=reasons1, y=count1)) +
  geom_bar(stat="identity", col = "skyblue", fill = "skyblue")+
  geom_text(aes(label=count1), vjust=1.6, color="white", size=3.5)+
  theme_bw()+
  ggtitle("Reasons that cause the patients who fail to place the referrals")
```

\newpage

#### 2.Reasons that cause the patients fail to schedule genetic meetings
\
\
We next look into the reasons that lead the patients to not to schedule genetic meetings. In the bar plot shown below, the main reason that cause patients to drop out is 'no response by call/mail'. There are 9 patients (about 27%) who did not schedule the meetings due to this reason. The rest patients who fail to schedule genetic meetings are mainly caused by 'not addressed' (15%), 'passed' (15%'), and 'pt declined' (18%). 


```{r echo=FALSE, warning=FALSE, message=FALSE, fig.cap = 'Reasons for Patients without Genetic Consouling Meetings Scheduled'}
df2 <- dat[c(18,19)]

df2$pt <- ifelse(str_detect(df2[,2], pattern = 'pt'), 1, 0)
df2$no <- ifelse(str_detect(df2[,2], pattern = 'no'), 1, 0)
df2$neverc <- ifelse(str_detect(df2[,2], pattern = 'never contacted'), 1, 0)
df2$nevers <- ifelse(str_detect(df2[,2], pattern = 'never schd'), 1, 0)
df2$address <- ifelse(str_detect(df2[,2], pattern = 'address'), 1, 0)
df2$pass <- ifelse(str_detect(df2[,2], pattern = 'pass'), 1, 0)

count2 <- c(sum(df2[3]=="1")-sum(df2[8]=="1"),sum(df2[4]=="1")-7,sum(df2[5]=="1"),
            sum(df2[6]=="1"),sum(df2[7]=="1"),sum(df2[8]=="1"))
reasons2 <- c("pt deferred/ declined ","no responses","never contancted", 
              "never schd", "not addressed", "passed")
result2<- data.frame(reasons2,count2)

ggplot(data=result2, aes(x=reasons2, y=count2)) +
  geom_bar(stat="identity", col = "pink", fill = "pink")+
  geom_text(aes(label=count2), vjust=1.6, color="white", size=3.5)+
  theme_bw() +
  ggtitle("Reasons that cause the patients fail to schedule genetic meetings")
```


\newpage

## Appendix

In the appendix part, we include the code about how we calculate the ratio above. And we also include the encoding method for Figure 6 and Figure 7. The additional stacked bar plots are also attached in this part.

### 1. Encoding Method for Figure 6
Encoding: 
confirm: confirm dx then chose care at ECMC, workup/confirm dx;

trial: phase III trial, trial screen;

palliative: palliative/ supportive, recommended supportive/ palliative care, other - palliative/ supportive;

testing: other - prior testing completed at Roswell (2017), other - prior testing previously completed outside;

hospice: hospice, then inpatient then enrolled in hospice, other - hospice


### 2. Encoding Method for Figure 7
Encoding: 
pt declined: pt deferred/not interested, pt deferred/ declined, other - pt declined x2;

No response/mail: other - no response to calls/VM or mailed letter, other - no call back, inpatient to inpatient hospice;

Never schd: no never schd, never seen;

Not addressed: referral not addressed - active;

Never contacted: never contacted - d/c via pt discharge;

passed: pt passed before referral reviewed, pt passed before appt offered, never contacted - d/c via pt discharge;


```{r include = FALSE}
#overall referral rate
sum(dat_1$test)/length(dat_1$test)

#referral rate before and after two options
#before
sum(dat_1$test[dat_1$`Initial Appt (#1-30)`<=22.2])/length(dat_1$test[dat_1$`Initial Appt (#1-30)`<=22.2])
#after
sum(dat$test[dat_1$`Initial Appt (#1-30)`>22.2])/length(dat_1$test[dat_1$`Initial Appt (#1-30)`>22.2])

```

```{r include = FALSE}
#before and after two options

nrow(dat_1 %>% filter(`Initial Appt (#1-30)`<= 22.2)) #total patients before
nrow(dat_1 %>% filter(`Initial Appt (#1-30)` > 22.2)) #total patients after

nrow(dat_1 %>% filter(`Initial Appt (#1-30)`<= 22.2& `# Genetics Referrals`>=1)) #total patients who had referrals before 
nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2& `# Genetics Referrals`>=1) ) #total patients who had referrals after


nrow(dat_1 %>% filter(`Initial Appt (#1-30)`<= 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'))#total patients who scheduled the appt before 
nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'))#total patients who scheduled the appt after

nrow(dat_1 %>% filter(`Initial Appt (#1-30)`<= 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'))#total patients who scheduled the appt before
nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'))#total patients who scheduled the appt after

nrow(dat_1 %>% filter(`Initial Appt (#1-30)`<= 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'& `Testing Ordered` == 'yes'))#total patients who ordered the test before 
nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'& `Testing Ordered` == 'yes'))#total patients who ordered the test after 


nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Virtual Offered` %in% c('yes - preference', 'yes - preference (distance)', 'yes'))) # patients who prefered virtual and scheduled the test
nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Virtual Offered` %in% c('no', 'no - preference'))) # patients who prefered person and scheduled the test

table(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes') %>% select(`Virtual Offered`))

nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes' & `Virtual Offered` %in% c('yes - preference', 'yes - preference (distance)', 'yes'))) # patients who prefered virtual and attended the test
nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'& `Virtual Offered` %in% c('no', 'no - preference'))) # patients who prefered person and attended the test

table(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes') %>% select(`Virtual Offered`))


table(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'&`Testing Ordered` == 'yes') %>% select(`Virtual Offered`))

```

### 3. More stacked bar plots

Figure.8 contains both of acceptance and rejection groups of patients who ordered testing diagnosed with different stages of cancer. The numbers and percentages on bars are same with the previous corresponded bar plots. And Figure.9 presents the relationship among patients in diversity of ECOG status and the referral rates in acceptance and rejection groups.

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.cap = 'Proportion of Patients Who Ordered Tests in each Cancer Stage', fig.height = 7, fig.width= 6}
p4 <- ggplot(df5, aes(x = `Stage at Dx (#0-4)`, y = n, fill = `Testing Ordered`)) +
   geom_bar(position = position_stack(), stat = "identity") +
   geom_text(aes(label = label), position = position_stack(vjust = 0.5), size = 3, color="white")+
   ylab("Frequency")+
   xlab("stages")

p3 <- ggplot(df5, aes(x = `Stage at Dx (#0-4)`, y = n, fill = `Testing Ordered`)) +
   geom_bar(position = "dodge", stat = "identity") +
   geom_text(aes(label=n), position=position_dodge(width=0.9), vjust=-0.25)+
   ylab("Frequency")+
   xlab("stages")

ggarrange(p4 , p3, ncol = 1, nrow = 2)

```



```{r echo=FALSE, warning=FALSE, message=FALSE, fig.cap = 'Proportions of Patients Who Ordered Tests in each ECOG Level', fig.height = 7, fig.width= 6}


p1 <- ggplot(df6, aes(x = `ECOG at Initial (#0-4)`, y = n, fill = `Testing Ordered`)) +
   geom_bar(position = position_stack(), stat = "identity") +
   geom_text(aes(label = label), position = position_stack(vjust = 0.5), size = 3, color="white")+
   ylab("Frequency")+
   xlab("ECOG")

p2 <- ggplot(df6, aes(x = `ECOG at Initial (#0-4)`, y = n, fill = `Testing Ordered`)) +
   geom_bar(position = "dodge", stat = "identity") +
   geom_text(aes(label=n), position=position_dodge(width=0.9), vjust=-0.25)+
   ylab("Frequency")+
   xlab("ECOG")

ggarrange(p1 , p2, ncol = 1, nrow = 2)
```

