---
title: "consutling"
author: "Zhihui Zhang 
         Shicong Wang"
date: "3/28/2022"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r , include = FALSE}
#package we use
library(readr)
library(stringr)
library(ggplot2)
library(dplyr)
library(stringr)
```


```{r , include = FALSE}
#import data
dat <- read_csv('data.csv')
```

## Introduction
The main purpose of this project is evaluating the genetic counseling and germline genetic testing process within the pancreatic adenocarcinoma patient population at Roswell Park Comprehensive Cancer Center. We are mainly interested in investigating the reasons that patients opt out of genetic testing and genetic counseling and the trend of referral rate after the guideline came out. The report can be divided into three parts. The first part will analyze the reason in each step on why people drop out. The second part will focus on the overall trend of the referral rate for 30 months. The third part will compare the referral rate before and after the genetic center offered telehealth and help to answer the question whether people prefer telehealth to in-person meeting. 


## Data Processing 

```{r , include = FALSE}
#second opinion tag 
dat$second_opinion <- str_detect(dat$`Reason No Referral`, pattern = 'second opinion') | str_detect(dat$`Somatic Testing
(y,n,other)`, pattern = 'second opinion')

#number of patients with second opinion
sum(dat$second_opinion)

#the data we will use for the further analysis(exclude patients with second opinion)
dat_1 <- dat[dat$second_opinion == FALSE,]

#referral rate = ordered test / total number
#re-encoding 
dat_1$test <- ifelse(str_detect(dat_1$`Testing Ordered`, pattern = 'yes'), 1, 0)

#referral rate through time 
dat_1$`Initial Appt (#1-30)`[dat_1$`Initial Appt (#1-30)`> 15 & dat_1$`Initial Appt (#1-30)`< 16]  <- 15
dat_1$`Initial Appt (#1-30)`[dat_1$`Initial Appt (#1-30)`> 22 & dat_1$`Initial Appt (#1-30)`< 23]  <- 22

```


## I.Referral rate and its analysis

The overall referral rate is 0.3548387. There are 55 patients ordered the germline genetic test among 155 patients. We will divide the referral process into four steps and analyze the reasons of dropping of during each step.  

### Part.I Genetic referrals placed

```{r,include = FALSE}
data<- dat_1

# probability that place genetic referrals
data$`Genetics Referrals`<- ifelse(data$`# Genetics Referrals` >0, "yes", "no")
sum(data$`Genetics Referrals` == "yes") / length(data$`Genetics Referrals`)
sum(data$`Genetics Referrals` == "no") / length(data$`Genetics Referrals`)

```

|Options          |Number of patients|  Ratio |
|----------------:|-----------------:|-------:|
|Yes - Acceptance |               111|   0.716| 
|No               |                44|   0.284|  
|Total            |               155|   -    | 

### Part II. Genetic meeting scheduled
```{r,include = FALSE}
# probability that schedule genetic meeting
data$`Genetics Appt Scheduled`<- ifelse(data$`Genetics Appt Scheduled` == "yes", "yes" , "no")
data$`Genetics Appt Scheduled2`<- ifelse(data$`Genetics Appt Scheduled` == "yes", "yes" ,
                                      ifelse(data$`Genetics Appt Scheduled` =="no", "no", "other"))
sum(data$`Genetics Appt Scheduled` == "yes") / sum(data$`Genetics Referrals` == "yes")
```

|Options          |Number of patients|  Ratio|
|----------------:|-----------------:|------:|
|Yes - Acceptance |                81|   0.73| 
|No               |                30|   0.27|  
|Total            |               111|   -   | 

### Part III. Genetic meeting attended
```{r,include = FALSE}
data$`Attended Appt`[data$`Attended Appt` == "yes x2"] <- "yes"
data$`Attended Appt`[data$`Attended Appt` == "???" | data$`Attended Appt` == "n/a"] <- "NA"

data$`Attended Appt`<- ifelse(data$`Attended Appt` == "yes", "yes", "no")
data$`Attended Appt2`<- ifelse(data$`Attended Appt` == "yes", "yes", 
                            ifelse(data$`Attended Appt` == "no", "no", "other"))
# probability that attend the appointment
# sum(data$`Attended Appt` %in% 'yes') / sum(data$`Genetics Appt Scheduled` %in% 'yes')

```

|Options          |Number of patients|  Ratio|
|----------------:|-----------------:|------:|
|Yes - Acceptance |                59|   0.73| 
|No               |                22|   0.27|  
|Total            |                81|   -   | 

### Part IV. Genetic counseling ordered 
```{r,include = FALSE}
data$`Testing Ordered`[data$`Testing Ordered` =="yes on second"]<- "yes"
data$`Testing Ordered`[data$`Testing Ordered` == "NO - PREVIOUSLY COMPLETED OUTSIDE" | data$`Testing Ordered` == "other - discussed DNA banking b/c hospice"]<- "no"
data$`Testing Ordered`[data$`Testing Ordered` =="??" | data$`Testing Ordered` =="n/a" ]<- "na"
data$`Testing Ordered`<- ifelse(data$`Testing Ordered` == "yes", "yes", "no")
data$`Testing Ordered2`<- ifelse(data$`Testing Ordered` == "yes", "yes", 
                               ifelse(data$`Testing Ordered` == "no","no","other"))
# probability that order genetic counseling 
# sum(data$`Testing Ordered` %in% 'yes') / sum(data$`Attended Appt` %in% 'yes')
```

|Options          |Number of patients|  Ratio|
|----------------:|-----------------:|------:|
|Yes - Acceptance |                55|   0.93| 
|No               |                 4|   0.07|  
|Total            |                59|   -   | 


### V. Proportions of individuals who proceed through each step of the process:
The following Sankey plot illustrates how many individuals succeed (currently labeled “Accept”) at each step of the genetic counseling process. 

```{r,include = FALSE}
#library(networkD3)
#source<- c("1st Step Accept","1st Step Accept","1st Step Reject","2nd Step Accept","2nd Step Accept","2nd Step Reject","3rd Step Accept","3rd Step Accept","3rd Step Reject","4th Step Reject")
#target<- c("2nd Step Accept","2nd Step Reject","Rejection 1","3rd Step Accept","3rd Step Reject","Rejection 2","4th Step Accept","4th Step Reject","Rejection 3","Rejection 4")
#value<- c(81,30,44,59,22,30,55,4,22,4)

#link<- data.frame(source,target,value)

#node <- data.frame(
 #name=c(as.character(link$source), as.character(link$target)) %>% 
    #unique()
  #)

#link$IDsource <- match(link$source, node$name)-1 
#link$IDtarget <- match(link$target, node$name)-1
 
#p2 <- sankeyNetwork(Links = link, Nodes = node,
                     #Source = "IDsource", Target = "IDtarget",
                     #Value = "value", NodeID = "name", 
                     #sinksRight=FALSE,fontSize=18)

#p2
```

## II. Other Analysis

### I.Cancer Stage and Genetic Counseling Referral Rate

We thought that the stage of cancer might explain some of the variation in referral rate. The following stacked bar plot is an early attempt to visualize the proportions of individuals who ordered testing given that they were diagnosed with certain stages of cancer. Currently, we don’t see much of a pattern with the proportion of testing ordered and the stage, though this may be due to the fact that so many of the patients are diagnosed at stage 4 (the smaller proportions may be unreliable). If the client is interested in us exploring this, we can try other ways of visualizing this (such as adding error bars). 

```{r echo=FALSE, warning=FALSE, message=FALSE}
data$`Testing Ordered` <- ifelse(str_detect(data$`Testing Ordered`, 'yes'), 'yes', 'no')
df5<- data %>%
  count(`Testing Ordered`,`Stage at Dx (#0-4)`) %>%
  arrange(`Testing Ordered`, desc(`Stage at Dx (#0-4)`)) %>% # Rearranging in stacking order  
  group_by(`Stage at Dx (#0-4)`) %>%
  mutate(Freq2 = cumsum(n), # Calculating position of stacked Freq
         prop = 100*n/sum(n))  # Calculating proportion of Freq

df5$label = paste0(sprintf("%.2f", df5$prop), "%")

ggplot(df5, aes(x = `Stage at Dx (#0-4)`, y = n, fill = `Testing Ordered`)) +
   geom_bar(position = position_stack(), stat = "identity") +
   geom_text(aes(label = label), position = position_stack(vjust = 0.5), size = 3, color="white")+
   ylab("Frequency")+
   xlab("Stages")
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(df5, aes(x = `Stage at Dx (#0-4)`, y = n, fill = `Testing Ordered`)) +
   geom_bar(position = "dodge", stat = "identity") +
   geom_text(aes(label=n), position=position_dodge(width=0.9), vjust=-0.25)+
   ylab("Frequency")+
   xlab("Stages")
```

### II. ECOG and Referral Rate

ECOG is an indicator of the wellness level of the patients. In the following section, we want to investigate the relationship between ECOG of patinets and the referral rate

```{r echo=FALSE, warning=FALSE, message=FALSE}

df6<- data %>%
  count(`Testing Ordered`,`ECOG at Initial (#0-4)`) %>%
  arrange(`Testing Ordered`, desc(`ECOG at Initial (#0-4)`)) %>% # Rearranging in stacking order  
  group_by(`ECOG at Initial (#0-4)`) %>%
  mutate(Freq2 = cumsum(n), # Calculating position of stacked Freq
         prop = 100*n/sum(n))  # Calculating proportion of Freq

df6$label = paste0(sprintf("%.2f", df6$prop), "%")

ggplot(df6, aes(x = `ECOG at Initial (#0-4)`, y = n, fill = `Testing Ordered`)) +
   geom_bar(position = position_stack(), stat = "identity") +
   geom_text(aes(label = label), position = position_stack(vjust = 0.5), size = 3, color="white")+
   ylab("Frequency")+
   xlab("ECOG")
```

```{r echo = FALSE, warning=FALSE, message=FALSE}
ggplot(df6, aes(x = `ECOG at Initial (#0-4)`, y = n, fill = `Testing Ordered`)) +
   geom_bar(position = "dodge", stat = "identity") +
   geom_text(aes(label=n), position=position_dodge(width=0.9), vjust=-0.25)+
   ylab("Frequency")+
   xlab("ECOG")
```


## III. Change of referral rate

In this section, we visualize the trends of referral rate using both 3 months and 6 months as a period. We can learn whether the referral rate is increasing or not through the time. 

```{r, include = FALSE}
#referral rate through time 
dat_2 <- dat_1
dat_2$`Initial Appt (#1-30)`[dat_2$`Initial Appt (#1-30)`> 15 & dat_2$`Initial Appt (#1-30)`< 16]  <- 15
dat_2$`Initial Appt (#1-30)`[dat_2$`Initial Appt (#1-30)`> 22 & dat_2$`Initial Appt (#1-30)`< 23]  <- 22

dat_2$cut_3 <- cut(dat_2$`Initial Appt (#1-30)`, c(1, seq(3, 30, by=3)), include.lowest = T) 
dat_2$cut_6 <- cut(dat_2$`Initial Appt (#1-30)`, c(1, seq(6, 30, by=6)), include.lowest = T) 

```


```{r echo=FALSE, warning=FALSE, message=FALSE}
dat1 <- dat_2 %>% group_by(`Initial Appt (#1-30)`,cut_3) %>%
  summarize(referral_rate= sum(test)/ n(), cut_3) %>% distinct()

dat2 <- dat1 %>% group_by(cut_3) %>%
 summarise(sd = sd(referral_rate),
          referral_rate = mean(referral_rate))


ggplot(dat1 , aes(x = cut_3, y = referral_rate)) +
  geom_jitter(position = position_jitter(0.15), color = "darkgray") + 
  geom_line(aes(group = 1), color="#69b3a2", data = dat2) +
  geom_errorbar(data = dat2, aes(ymin = referral_rate-sd, ymax =referral_rate+sd), width = 0.2)  +
  geom_point(data = dat2, size = 2) + 
  labs(title="Referral rate change per 3 month", x="Month", y="Referral rate")
```
The plot above shows us the trend of referral rate using 3 month as a period.We jitter the points to avoid overlaps. The plot above does not indicate there is any increasing or decreasing trend. 

```{r echo=FALSE, message=FALSE}
dat_2$cut_6 <- cut(dat_2$`Initial Appt (#1-30)`, c(1, seq(6, 30, by=6)), include.lowest = T) 

dat1 <- dat_2 %>% group_by(`Initial Appt (#1-30)`,cut_6) %>%
  summarize(referral_rate= sum(test)/ n(), cut_6) %>% distinct()

dat2 <- dat1 %>% group_by(cut_6) %>%
 summarise(sd = sd(referral_rate),
          referral_rate = mean(referral_rate))


ggplot(dat1 , aes(x = cut_6, y = referral_rate)) +
  geom_jitter(position = position_jitter(0.2), color = "darkgray") + 
  geom_line(aes(group = 1), data = dat2,  color="#69b3a2") +
  geom_errorbar(data = dat2, aes(ymin = referral_rate-sd, ymax =referral_rate+sd), width = 0.2)  +
  geom_point(data = dat2, size = 2)+ 
   labs(title="Referral rate change per 6 month", x="Month", y="Referral rate")
```

The visualization above tells us the change of the referral rate per 6 month. Since there are a lot of variations of the referral rate in each time interval, we can not conclude that there is an obvious increasing relationship between time and referral rate. 



\newpage

## III.Virtual/Teleheath Impact

In this part, we want to investigate if there is any difference of the referral rate before and after telehealth is provided as an preference options for the patients. 

|Period|Total number of patients|Number of patients who placed a referral|Number of Patients who scheduled a meeting|Number of patients who attended the meeting|Number of Patients who ordered the test|Referral rate|
|----------:|----------:|----------:|----------:|----------:|----------:|----------:|
|Before     |        120|         82|         55|         40|         37|     0.3083|
|After      |         35|         29|         26|         18|         17|     0.4857|
| Telehealth|     -     |     -     |         17|         14|         13|     -     |
|  In-person|     -     |     -     |          5|          4|          4|     -     |
|         NA|     -     |     -     |          4|          0|          0|     -     |
|Total      |        155|        111|         81|         58|         54|     0.3484| 

Qeustion: 


\newpage

## IV.Analysis on second opinion patients


Among 65 patients who had second opinion in the data, 3 of them still ordered the germline genetic testing.

```{r, include = FALSE}
dat_3 <- dat[dat$second_opinion == TRUE, ]
nrow(dat_3)
nrow(dat_3 %>% filter(`# Genetics Referrals`>=1))
nrow(dat_3 %>% filter(`# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'))
nrow(dat_3 %>% filter(`# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'))
sum(dat_3$`Testing Ordered` == 'yes')
```
  

|Period|Total number of patients|Number of patients who placed a referral|Number of Patients who scheduled a meeting|Number of patients who attended the meeting|Number of Patients who ordered the test|Referral rate|
|----------:|----------:|----------:|----------:|----------:|----------:|----------:|
|Before     |         65|         11|          8|          6|          3|    0.04615|






\newpage

## Appendix
```{r include = FALSE}
#overall referral rate
sum(dat_1$test)/length(dat_1$test)

#referral rate before and after two options
#before
sum(dat_1$test[dat_1$`Initial Appt (#1-30)`<=22.2])/length(dat_1$test[dat_1$`Initial Appt (#1-30)`<=22.2])
#after
sum(dat$test[dat_1$`Initial Appt (#1-30)`>22.2])/length(dat_1$test[dat_1$`Initial Appt (#1-30)`>22.2])

```

```{r include = FALSE}
#before and after two options

nrow(dat_1 %>% filter(`Initial Appt (#1-30)`<= 22.2)) #total patients before
nrow(dat_1 %>% filter(`Initial Appt (#1-30)` > 22.2)) #total patients after

nrow(dat_1 %>% filter(`Initial Appt (#1-30)`<= 22.2& `# Genetics Referrals`>=1)) #total patients who had referrals before 
nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2& `# Genetics Referrals`>=1) ) #total patients who had referrals after


nrow(dat_1 %>% filter(`Initial Appt (#1-30)`<= 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'))#total patients who scheduled the appt before 
nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'))#total patients who scheduled the appt after

nrow(dat_1 %>% filter(`Initial Appt (#1-30)`<= 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'))#total patients who scheduled the appt before
nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'))#total patients who scheduled the appt after

nrow(dat_1 %>% filter(`Initial Appt (#1-30)`<= 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'& `Testing Ordered` == 'yes'))#total patients who ordered the test before 
nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'& `Testing Ordered` == 'yes'))#total patients who ordered the test after 


nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Virtual Offered` %in% c('yes - preference', 'yes - preference (distance)', 'yes'))) # patients who prefered virtual and scheduled the test
nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Virtual Offered` %in% c('no', 'no - preference'))) # patients who prefered person and scheduled the test

table(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes') %>% select(`Virtual Offered`))

nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes' & `Virtual Offered` %in% c('yes - preference', 'yes - preference (distance)', 'yes'))) # patients who prefered virtual and attended the test
nrow(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'& `Virtual Offered` %in% c('no', 'no - preference'))) # patients who prefered person and attended the test

table(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes') %>% select(`Virtual Offered`))


table(dat_1 %>% filter(`Initial Appt (#1-30)`> 22.2 & `# Genetics Referrals`>=1&`Genetics Appt Scheduled`== 'yes'& `Attended Appt` == 'yes'&`Testing Ordered` == 'yes') %>% select(`Virtual Offered`))


```
